---
title: "X is the new Y - contrastive"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    collapsed: false
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

session info:

```{r sessionInfo, message = FALSE, warning = FALSE}

sessionInfo()

```

Install and load packages

```{r pkg, message = FALSE, warning = FALSE}

# install CRAN packages (if not yet installed)
sapply(c("readxl", "tidyverse", "patchwork", "devtools", "vroom", "DT"), function(x) if(!is.element(x, installed.packages())) install.packages(x, dependencies = T))

# install non-CRAN packages (if not yet installed)
if(!is.element("concordances", installed.packages())) {
devtools::install_github("hartmast/concordances")
}

# if this doesn't work, check sfla.ch for the package
if(!is.element("collostructions", installed.packages())) {
  install.packages("https://sfla.ch/wp-content/uploads/2021/02/collostructions_0.2.0.tar.gz", repos = NULL)
}

# load packages
library(readxl)
library(tidyverse)
library(patchwork)
library(vroom)
library(collostructions)
library(DT)

```

## Data wrangling

```{r}

# English
en <- read_xlsx("ENCOW_x_is_the_new_y_without_false_hits.xlsx")

# Spanish
es <- read_xlsx("ESCOW_x_nuev_y_without_false_hits.xlsx")

# German
de <- read_xlsx("DECOW_x_ist_das_neue_y_without_false_hits.xlsx")


```

omit uncertain hits

```{r}
en <- filter(en, keep=="y")
es <- filter(es, keep=="y")
de <- filter(de, keep=="y")

```

## Frequency and productivity measures

Get types, tokens, and hapax legomena for the three languages.

```{r}

# add lowercase lemmas
es$lemma_x <- tolower(es$Lemma_x)
es$lemma_y <- tolower(es$Lemma_y)

de$lemma_x <- tolower(de$Lemma_x)
de$lemma_y <- tolower(de$Lemma_y)

en$lemma_x <- tolower(en$Lemma_x)
en$lemma_y <- tolower(en$Lemma_y)

# add x-y combinations

es$lemma_x_y <- paste0(es$lemma_x, "-", es$lemma_y)
en$lemma_x_y <- paste0(en$lemma_x, "-", en$lemma_y)
de$lemma_x_y <- paste0(de$lemma_x, "-", de$lemma_y)

# get hapaxes --> add frequency counts
en <- en %>% group_by(lemma_x) %>% 
  add_count(name = "count_x") %>%
  ungroup %>% group_by(lemma_y) %>%
  add_count(name = "count_y") %>%
  ungroup %>% group_by(lemma_x_y) %>%
  add_count(name = "count_x_y")

es <- es %>% group_by(lemma_x) %>% 
  add_count(name = "count_x") %>%
  ungroup %>% group_by(lemma_y) %>%
  add_count(name = "count_y") %>%
  ungroup %>% group_by(lemma_x_y) %>%
  add_count(name = "count_x_y")

de <- de %>% group_by(lemma_x) %>% 
  add_count(name = "count_x") %>%
  ungroup %>% group_by(lemma_y) %>%
  add_count(name = "count_y") %>%
  ungroup %>% group_by(lemma_x_y) %>%
  add_count(name = "count_x_y")

# table
tibble(languages = c("English", "German", "Spanish"),
       tokens = c(nrow(en), nrow(de), nrow(es)),
       types_x = c(length(unique(en$lemma_x)),
                   length(unique(de$lemma_x)),
                   length(unique(es$lemma_x))),
       types_y = c(length(unique(en$lemma_y)),
                   length(unique(de$lemma_y)),
                   length(unique(es$lemma_y))),
       types_x_y = c(length(unique(en$lemma_x_y)),
                   length(unique(de$lemma_x_y)),
                   length(unique(es$lemma_x_y))),
       hapaxes_x = c(length(which(en$count_x==1)),
                     length(which(de$count_x==1)),
                     length(which(es$count_x==1))),
       hapaxes_y = c(length(which(en$count_y==1)),
                     length(which(de$count_y==1)),
                     length(which(es$count_y==1))),
       hapaxes_x_y = c(length(which(en$count_x_y==1)),
                     length(which(de$count_x_y==1)),
                     length(which(es$count_x_y==1)))
       
       ) 



```


## Collostructional analysis

For the collostructional analysis, we read in the lemma lists for the English, German, and Spanish data.

```{r, message=FALSE, eval=FALSE}

decow <- vroom("decow16bx.lp.tsv.gz", col_names = c("Lemma", "POS", "Freq"))
escow <- vroom("escow14ax.freq0.l.zip")
encow <- vroom("encow16ax.lp.tsv.gz", col_names = c("Lemma", "POS", "Freq"))

```

Find the lemmas that actually occur in the datasets

```{r, eval=FALSE}

# lemma list ES
es_lemmas <- unique(c(es$Lemma_x, es$Lemma_y))

# lemma list DE
de_lemmas <- unique(c(de$Lemma_x, de$Lemma_y))

# which ESCOW lemmas in es_lemmas?
find_lemmas <- which(escow$token... %in% es_lemmas)
escow <- escow[find_lemmas,]
write_csv(escow, "es_cow_lemmas.csv")

# which DECOW lemmas in de_lemmas?
find_lemmas_de <- which(decow$Lemma %in% de_lemmas)
decow <- decow[find_lemmas_de,]
write_csv(decow, "de_cow_lemmas.csv")


```

Read filtered frequency lists

```{r}

# ESCOW lemma frequencies
escow <- read_csv("es_cow_lemmas.csv")

# DECOW lemma frequencies
decow <- read_csv("de_cow_lemmas.csv")

# ESCOW pos frequencies
es_pos <- read_csv("frequency_lists/ESCOW_pos_frequencies.csv")

# DECOW pos frequencies
de_pos <- read_csv("frequency_lists/decow_pos_frequencies.csv")

# get total number of nouns and adjectives
n_a_freq <- filter(es_pos, pos %in% c("NOUN", "ADP", "PROPN", "ADJ")) %>% select(frequency) %>% sum

n_a_freq_de <- filter(de_pos, tag %in% c("NN", "ADJA", "ADJD", "NE", "NN"))

# collostructional analysis ----


# for the collostructional analysis,
# reduce two-word lemmas to the last word
# (compound heads for compounds, last names
# for person names)
de$lemma_x <- gsub(".* ", "", de$lemma_x)
de$lemma_y <- gsub(".* ", "", de$lemma_y)


# create input for coll. anal. for
# x slot
input_x <- table(es$lemma_x) %>% as_tibble(.name_repair="unique") %>% setNames(c("Lemma", "Freq_in_cxn"))

input_x_de <- table(de$lemma_x) %>% as_tibble(.name_repair="unique") %>% setNames(c("Lemma", "Freq_in_cxn"))

# lowercase lemma in escow
escow$lemma <- tolower(escow$token...)

# lowercase lemma in decow
decow$lemma <- tolower(decow$Lemma)

decow <- decow %>% group_by(lemma) %>% summarise(
  Freq = sum(Freq)
)

# join dataframes to combine lemma and corpus freqs
input_x <- left_join(input_x, select(escow, lemma, f_raw),
          by = c("Lemma" = "lemma"))


input_x_de <- left_join(input_x_de, select(decow, lemma, Freq),
          by = c("Lemma" = "lemma"))

# replace NA by 0
input_x <- input_x %>% replace_na(list(Freq_in_cxn = 0, f_raw = 0))

input_x_de <- input_x_de %>% replace_na(list(Freq_in_cxn = 0, Freq = 0))

# omit all with f_raw < Freq_in_cxn
input_x <- input_x[which(input_x$f_raw>=input_x$Freq_in_cxn),]

input_x_de <- input_x_de[which(input_x_de$Freq >= input_x_de$Freq_in_cxn),]

# same for y -----

# generate input for coll. anal.
input_y_de <- table(de$lemma_y) %>% as_tibble(.name_repair="unique") %>% setNames(c("Lemma", "Freq_in_cxn"))

# join dataframes to combine lemma and corpus freqs
input_y <- left_join(input_y_de, select(escow, lemma, f_raw),
          by = c("Lemma" = "lemma"))


input_y_de <- left_join(input_y_de, select(decow, lemma, Freq),
          by = c("Lemma" = "lemma"))

# replace NA by 0
input_y_de <- input_y_de %>% replace_na(list(Freq_in_cxn = 0, Freq = 0))

input_y_de <- input_y_de[which(input_y_de$Freq >= input_y_de$Freq_in_cxn),]


```

## Simple collexeme analysis - Spanish

```{r}

# coll. anal.
collex_x <- collex(as.data.frame(input_x), corpsize = n_a_freq)


# same for y slot -----

# create input for coll. anal. for
# x slot
input_y <- table(es$lemma_y) %>% as_tibble(.name_repair="unique") %>% setNames(c("Lemma", "Freq_in_cxn"))


input_y <- left_join(input_y, select(escow, lemma, f_raw),
          by = c("Lemma" = "lemma"))

# replace NA by 0
input_y <- input_y %>% replace_na(list(Freq_in_cxn = 0, f_raw = 0))

# omit all with f_raw < Freq_in_cxn
input_y <- input_y[which(input_y$f_raw>=input_y$Freq_in_cxn),]

# coll. anal.
collex_y <- collex(as.data.frame(input_y), corpsize = n_a_freq)


```

Results for x slot

```{r}
collex_x %>% DT::datatable()
```

Results for y slot

```{r}
collex_y %>% DT::datatable()
```

## Simple collexeme analysis - German

```{r}

collex_x_de <- input_x_de %>% as.data.frame %>% collex(corpsize = sum(n_a_freq_de$frequency))

collex_y_de <- input_y_de %>% as.data.frame %>% collex(corpsize = sum(n_a_freq_de$frequency))


```

Results x slot

```{r}

collex_x_de %>% DT::datatable()

```

Results y slot

```{r}

collex_y_de %>% DT::datatable()

```


## Dinstinctive collexeme analysis

```{r}

es %>% ungroup %>% select(lemma_x, lemma_y) %>% as.data.frame() %>% collex.covar() %>% DT::datatable()

de %>% ungroup %>% select(lemma_x, lemma_y) %>% as.data.frame() %>% collex.covar() %>% DT::datatable()

```
