---
title: "X is the new Y - contrastive"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    collapsed: false
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminaries

session info:

```{r sessionInfo, message = FALSE, warning = FALSE}

sessionInfo()

```

Install and load packages

```{r pkg, message = FALSE, warning = FALSE}

# install CRAN packages (if not yet installed)
sapply(c("readxl", "tidyverse", "patchwork", "devtools", "vroom", "DT"), function(x) if(!is.element(x, installed.packages())) install.packages(x, dependencies = T))

# install non-CRAN packages (if not yet installed)
if(!is.element("concordances", installed.packages())) {
devtools::install_github("hartmast/concordances")
}

# if this doesn't work, check sfla.ch for the package
if(!is.element("collostructions", installed.packages())) {
  install.packages("https://sfla.ch/wp-content/uploads/2021/02/collostructions_0.2.0.tar.gz", repos = NULL)
}

# load packages
library(readxl)
library(tidyverse)
library(patchwork)
library(vroom)
library(collostructions)
library(DT)

```

## Data wrangling

```{r}

# English
en <- read_xlsx("ENCOW_x_is_the_new_y_without_false_hits.xlsx")

# Spanish
es <- read_xlsx("ESCOW_x_nuev_y_without_false_hits.xlsx")

# German
de <- read_xlsx("DECOW_x_ist_das_neue_y_without_false_hits.xlsx")


```

omit uncertain hits

```{r}
en <- filter(en, keep=="y")
es <- filter(es, keep=="y")
de <- filter(de, keep=="y")

```

## Collostructional analysis

For the collostructional analysis, we read in the lemma lists for the English, German, and Spanish data.

```{r, message=FALSE, eval=FALSE}

decow <- vroom("/Users/stefanhartmann/Downloads/decow16bx.lp.tsv.gz", col_names = c("Lemma", "POS", "Freq"))
escow <- vroom("/Users/stefanhartmann/sciebo/Projekte/snowclones/ESCOW frequency lists/escow14ax.freq0.l.zip")
encow <- vroom("/Users/stefanhartmann/sciebo/Tutorials/collostructions_tutorial/data/encow16ax.lp.tsv.gz", col_names = c("Lemma", "POS", "Freq"))

```

Find the lemmas that actually occur in the datasets

```{r, eval=FALSE}

# lemma list ES
es_lemmas <- unique(c(es$Lemma_x, es$Lemma_y))

# lemma list DE
de_lemmas <- unique(c(de$Lemma_x, de$Lemma_y))

# which ESCOW lemmas in es_lemmas?
find_lemmas <- which(escow$token... %in% es_lemmas)
escow <- escow[find_lemmas,]
write_csv(escow, "es_cow_lemmas.csv")

# which DECOW lemmas in de_lemmas?
find_lemmas_de <- which(decow$Lemma %in% de_lemmas)
decow <- decow[find_lemmas_de,]
write_csv(decow, "de_cow_lemmas.csv")


```

Read filtered frequency lists

```{r}

# ESCOW lemma frequencies
escow <- read_csv("es_cow_lemmas.csv")

# DECOW lemma frequencies
decow <- read_csv("de_cow_lemmas.csv")

# ESCOW pos frequencies
es_pos <- read_csv("frequency_lists/ESCOW_pos_frequencies.csv")

# get total number of nouns and adjectives
n_a_freq <- filter(es_pos, pos %in% c("NOUN", "ADP", "PROPN", "ADJ")) %>% select(frequency) %>% sum

# collostructional analysis ----

# lowercase lemmas
es$lemma_x <- tolower(es$Lemma_x)
es$lemma_y <- tolower(es$Lemma_y)

# create input for coll. anal. for
# x slot
input_x <- table(es$lemma_x) %>% as_tibble(.name_repair="unique") %>% setNames(c("Lemma", "Freq_in_cxn"))

# lowercase lemma in escow
escow$lemma <- tolower(escow$token...)

input_x <- left_join(input_x, select(escow, lemma, f_raw),
          by = c("Lemma" = "lemma"))

# replace NA by 0
input_x <- input_x %>% replace_na(list(Freq_in_cxn = 0, f_raw = 0))

# omit all with f_raw < Freq_in_cxn
input_x <- input_x[which(input_x$f_raw>=input_x$Freq_in_cxn),]

# coll. anal.
collex_x <- collex(as.data.frame(input_x), corpsize = n_a_freq)


# same for y slot -----

# create input for coll. anal. for
# x slot
input_y <- table(es$lemma_y) %>% as_tibble(.name_repair="unique") %>% setNames(c("Lemma", "Freq_in_cxn"))


input_y <- left_join(input_y, select(escow, lemma, f_raw),
          by = c("Lemma" = "lemma"))

# replace NA by 0
input_y <- input_y %>% replace_na(list(Freq_in_cxn = 0, f_raw = 0))

# omit all with f_raw < Freq_in_cxn
input_y <- input_y[which(input_y$f_raw>=input_y$Freq_in_cxn),]

# coll. anal.
collex_y <- collex(as.data.frame(input_y), corpsize = n_a_freq)


```

Results for x slot

```{r}
collex_x %>% DT::datatable()
```

Results for y slot

```{r}
collex_y %>% DT::datatable()
```


## Dinstinctive collexeme analysis

```{r}

es %>% select(lemma_x, lemma_y) %>% as.data.frame() %>% collex.covar() %>% DT::datatable()

```
